---
title: "ROSSyndicate Project and Workflow Best Practices"
author: "ROSSyndicate"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
description: |
  The ROSSyndicate workflow best practices guide.
output: html_document
---

```{r, echo = F}
library(formatR)
library(bookdown)
library(servr)
library(tidyverse)
```

# About

This guide outlines the [ROSSyndicate's](https://rossyndicate.com/) project
management expectations and workflow best practices so that we can successfully
implement Radical Open Science.

## Overarching code/workflow philosphies of the ROSSyndicate

- We value literate, readable code without sacrificing reproducibility. This
means deploying understandable code and repositories that are accompanied by
narrative comments and text.We value transparency in our analyses and believe
that well documented repositories and code facilitate that.

- All code and software are licensed with the highly permissive MIT use license.
While we believe that all code and software should be public, we acknowledge
that other researchers and users may wish to keep products derived from our
code private. We intentionally avoid the GNU-GPL family of licenses that are
"viral" by nature, forcing all downstream repositories to also be fully open.
We do so because we acknowledge their are vulnerable populations that may want
to keep their code private for reasons we may not understand. We hope that this
encourages use of our code, and ultimately, creates an environment of sharing
with and empowering each other.


- GitHub contributions follow the 'fork' methodology that is common in open
science repositories (as opposed to the 'branch' method).

- All GitHub contributions are reviewed before being incorporated into the lab
organization repository, except upon initial repository set up.  These reviews
are completed internally through pull requests. We value 'bite-sized' pull
requests that usually only contain 500 lines of code.

- No code is formally published unless it has been consistently and intentionally
reviewed. We make an effort to include external review of our repositories
prior to formal publishing.

- We follow [the {tidyverse} style guide](https://style.tidyverse.org/). We use
the {styler} Addin in RStudio to assist in linting our code.

- We strive use a set of common styles specific to the look and feel of rendered
documents, which include: color palettes, ggthemes, and style.css files. These
are meant to reduce the amount of time that researchers in this lab spend on
relatively inconsequential decisions. Ideally the use of these aesthetics also
creates a common look and feel of all deliverables from the ROSSyndicate.

## Contents

This best practices guide serves as an introduction to Git and GitHub, the
lab's style and code philosophies, and the tools we use for project management.

## A living document

The workflow best practices document is a living document and all members of
the ROSSyndicate are welcome to submit edits to it via pull request.

## Origin

Adapted from the Geospatial Centroid's GitHub Workflow Document, Caitlin
Mothes, built out for ROSSyndicate by Katie Willi, B Steele, and Matt Brousil.

```{r, echo = F}
knitr::wrap_rmd('index.Rmd', width = 80, backup = NULL)
#note, this will not wrap text that are prefaced by any special characters (like bullets!)
```
